---
title: "Covid91 vaccine study Final2021"
author: "Adriana Lopez!"
date: "Spring 2021"
output: 
  html_document:
    number_sections: true
    toc: true

---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
OldData <-read.csv(file="covid91.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- OldData %>% mutate(infected = recode(infected, aids = "covid", none = "OK" )) %>% select(-one_of("X"))
```

# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups consisting of Males, Females, just the drug users, and just those in the LGBTQ community. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

# Methods 

The first step is isolating the subpopulations 
```{r}
maledata<- filter(FinalData,sex=="guy")
maledata$sex <- as.factor(maledata$sex)
```

# Males

## Bar Chart

```{r}
Males <- maledata %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))

# the group_by followed by summarize(count=n())

basicC <- ggplot(Males,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(Males,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This Bar Chart appears to show that out of approximately 340 people were infected with Covid-19 57% of the males who had Covid were given the drug, and about 43% were given the placebo. 

## Numerical Summaries

```{r}
table1 <- xtabs(~treatment + infected, data=maledata)
rowPerc(table1)
colPerc(table1)
print(table1)
```
The tables shows that about 340 males were infected with Covid-19 out of approximately 10,400 males in the study. The table shows that 3.72% of males who receive the drug got Covid, and 2.55% of the males who got the placebo contracted Covid. The table also shows, that males that got the vaccine were about 50% more likely to contract Covid-19, this suggest that the vaccine is not effective for males.

## Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
The Chi Squared table shows that the P-Value of 0.0006181 is sosmal  that it is highly unlikely that the results are by chance.  
$H_0$ the null hypothesis given taking the vaccine there is no difference of whether males can contract Covid-19 or not.  
$H_0$ the alternative hypothesis is that the vaccine in fact does make a difference in whether or not a male contracts Covid-19. Therefore we Reject the Null Hypothesis.

The odds ratio from the Fisher Exact test is 1.47:1, when taking the odds radio and computing it into a percent of effectiveness by using the formula 0/(1+0)*100 you get 59.5% effectiveness.  

# Females

# LGBTQ

# Drug Users

# Discussion & Conclusion


